First steps in synthetic algebraic geometry

### 2017-03-15 snippets/choice-site.tex
negneg-stability of open truth values

The Stacks Project then defines~$\Zar(S)$ as~$\Sh(\Sch_M/S)$, where~$\Sch_M/S$
is the small category of~$S$-schemes in~$M$, or equivalently
as~$\Sh(\Aff_M/S)$. This approach has the advantage that one doesn't have to
assume the existence of a Grothendieck universe.

>>>>

The Stacks Project then defines~$\Zar(S)$ as~$\Sh(\Sch_M/S)$, where~$\Sch_M/S$
is the small category of~$S$-schemes in~$M$, or equivalently
as~$\Sh(\Aff_M/S)$. This approach has the advantage that one doesn't have to
assume the existence of a Grothendieck universe; the \emph{partial
universe}~$M$ can be constructed entirely within ZFC set theory using
transfinite recursion.

### 2017-03-15 snippets/open-subfunctors.tex
negneg-stability of open truth values

\begin{prop}In~$\Zar(S)$, open truth values are~$\neg\neg$-stable: If~$U
\subseteq 1$ is an open truth value, and if~$U$ does \notnot hold, then~$U$
actually holds.\end{prop}

\begin{proof}Since~$U$ is an open truth value, there exists an ideal~$J
\subseteq \affl$ such that~$\affl/J$ is quasicoherent and such that~$U$ holds
if and only if~$1 \in J$. By assumption, the element~1 is \notnot an element
of~$J$; we want to verify that it's actually an element of~$J$.

Since~$\affl/J$ is quasicoherent, the canonical homomorphism
\[ \affl/J \longrightarrow XXX \]
XXX
\end{proof}

>>>>

In many flavours of synthetic topology, open truth values~$\varphi$
are~$\neg\neg$-stable in that~$\neg\neg\varphi$ implies~$\varphi$. With a
small caveat, this is true for open truth values in the big Zariski topos as
well.

\begin{prop}\label{prop:open-truth-values-stable}
Let~$U \hookrightarrow 1$ be a subfunctor in~$\Zar(S)$ such
that~$\Zar(S) \models \neg\neg U$. Then in any of the following
situations it follows that~$\Zar(S) \models U$:
\begin{enumerate}
\item $U$ is a quasicompact open truth value.
\item $U$ is an arbitrary open truth value and the site used to
define~$\Zar(S)$ is closed under domains of closed immersions.
\end{enumerate}
\end{prop}

\begin{proof}We give two proofs, an internal one and an external one, since
they pursue different ideas.

\emph{Internal proof.} Since~$U$ is an open truth value, there exists an ideal~$J
\subseteq \affl$ such that~$\affl/J$ is quasicoherent and such that~$U$ holds
if and only if~$1 \in J$. By assumption, the element~1 is \notnot an element
of~$J$; we want to verify that it's actually an element of~$J$.

By Proposition~\ref{prop:XXX}, the canonical homomorphism
\[ \affl/J \longrightarrow [\Spec(\affl/J), \affl] \]
is bijective; the assumptions of that proposition are satisfied in either of
the two situations. The set~$\Spec(\affl/J)$ is isomorphic to~$\brak{J = (0)}$.
Since~$\neg\neg(1 \in J)$, we also have~$\neg(J = (0))$.
Therefore~$\Spec(\affl/J)$ is empty and the codomain of the displayed
isomorphism is the zero algebra. Thus~$\affl/J$ is trivial as well, showing~$1
\in J$.

\emph{External proof.} Since~$U \hookrightarrow 1$ is an open subfunctor, there
is an open subscheme~$S_0 \subseteq S$ such that a morphism~$f : T \to S$
factors over~$S_0$ if and only if~$U(T)$ is inhabited. In both situations it's
possible to endow~$X \defeq S \subseteq S_0$ with the structure of a closed
subscheme such that~$X$ is contained in the site used to define~$\Zar(S)$.
By the universal property of~$S_0$, we have~$X \models \neg U$. Since~$\Zar(S)
\models \neg\neg U$, it follows that~$X$ is empty. Therefore~$S_0 = S$ and~$U$
is globally inhabited.
\end{proof}

\begin{cor}Let~$\gamma : \Delta \to X$ be a morphism in~$\Zar(S)$.
Let~$U \hookrightarrow X$ be an open subfunctor such that~$\Zar(S) \models
\gamma(0) \in U$. Then, in any of the situations in
Proposition~\ref{prop:open-truth-values-stable}, the morphism~$\gamma$ factors
over~$U$.
\end{cor}

\begin{proof}We give an internal proof. Let~$\varepsilon \in \Delta$.
Then~$\neg\neg(\varepsilon = 0)$. Therefore~$\neg\neg(\gamma(\varepsilon) \in
U)$. Since being an element of~$U$ is~$\neg\neg$-stable, it follows
that~$\gamma(\varepsilon) \in U$.
\end{proof}
